Highly Available Transactions: Virtues and Limitations

### 2013-03-22 writeup/evaluation.tex
meh; sort of crappy TPC-C writeup

Integrity constraints:
--Mostly SUM, AVG, COUNT
--This is okay and foreign key constraints are provided by TA

New Order:
--Lookup of warehouse codes is fine, are not mutated
--Assigning a version order is not okay, but we can still get uniqueness
--Stock adjustment is monotonic and can't go below zero due to replenish. However, might overstock!

Payment:
--Monotonic, simply increase the number of each respective field

Delivery:
--Problematic; is non-monotonic, removing from the order table
--Without additional semantics, have to block
--CALM analysis would catch this, enforce coordination
--Does \textit{not} block new New Order transactions

Order Status:
--Non-monotonic
--Read your writes may be problematic

While TPC-C isolation tests are for lock-based models (and HATs are,
by requirement, non-locking), consider TPC-E tests. Modifying the
description and image together is fine, etc.

>>>>

TPC-C consists of five transactions, capturing operation of a
real-world warehouse, including sales, payments, and deliveries. We
analyze each transaction and its interactions below.

\begin{itemize}
\item \textbf{New-Order} places an order for a variable quantity of
  data items, updating warehouse stock as needed. The New-Order
  transaction selects a sales district, assigns the new order an order
  ID number, adjusts the remaining warehouse stock, and writes a
  placeholder entry for the pending order. There are two mutations in
  this transaction which we must be careful of.

  First, assigning an order ID number is challenging. We can easily
  assign a \textit{unique} order ID, by, say, concatenating the user
  ID with a timestamp and server ID. However, the TPC-C specification
  requires that order numbers be sequentially assigned, which we
  cannot guarantee in HATs due to Lost Update. Accordingly, while HATs
  cannot provide compliant TPC-C execution due to a non-monotonic
  counter increase, they can allow placement of new orders, which is a
  commutative and monotonic operation

  Second, decrementing inventory counts is a potentially risky
  operation as, without coordination, inventory counts may become
  negative. However, TPC-C automatically ``refills'' inventory counts
  (by 91 items) when there are 10 or fewer items
  remaining. Accordingly, even in the presence of concurrent new
  orders to the same item, the stock will never fall below
  zero. However, if the system propagates ``refill'' operations
  between replicas, then an $R$ replica system might end up with
  $91(R-1)$ more stock than a single-site implementation.

\item \textbf{Payment} updates running balances for warehouses,
  districts, and customer records as well as providing an audit
  trail. All balance operations can be made commutative, so the
  transaction can execute under HATs.

\item \textbf{Order-Status} is a read-only transaction and can be
  safely executed under HATs. The TPC-C specification does not require
  recency bounds but, in practice, they may be useful. If a client is
  sticky with her replica set, she will read her own writes, which may
  be sufficient.

\item \textbf{Delivery} represents the fulfillment of a pending order
  created by a New Order transaction: it deletes the order from the
  pending list, updates the customer's balance, updates the order's
  carrier ID and delivery time, and updates the customer balance. The
  challenge in delivery comes in deletion, which is non-monotonic: an
  item might be removed twice from the pending table. If delivery can
  be made idempotent---that is, if multiple Delivery transactions have
  the same effect as one---then multiple clients can execute the same
  Deliveries without fatal consequences.

  In practice, given that a physical shipment is likely only delivered
  once, it is unlikely that multiple agents will attempt a Delivery
  transaction for the same order simultaneously. However, in this
  case, we have simply moved the non-monotonicity to the real world:
  the carrier that picks up the package for an order must ensure that
  she is the only carrier who has done so. While we can program around
  the other operations in the order, HATs cannot--without the help of
  business-specific processes--serve Delivery transactions. That said,
  New Orders may be placed.

\item \textbf{Stock-Level} is another read-only transaction and can be
  executed without any non-HAT guarantees, with the same caveats
  regarding recency.
\end{itemize}

Throughout execution, TPC-C requires the maintainence of several
integrity constraints. For example, Consistency Condition 1 (3.3.2.1)
requires that each warehouse's sales count must reflect the sum of its
subordinate sales districts. This integrity constraint spans two
tables but, given the ability to update both tables atomically via
transactional atomicity, can be easily maintained. Consistency
Conditions 4 through 12 (3.3.2.4-12) can similarly be satisfied by
applying updates atomically across tables. Consistency Conditions 2
and 3 (3.3.2.2-3) concern order ID assignment and are more
problematic: effectively, each New Order placed by a customer in a
district must have an ID that is one the prior order in that
district. This requires preventing Lost Update within each district's
table. However, as we have seen, the required transactions can execute
without these two conditions.

In summary, many---but not all---TPC-C transactions are well-served by
HATs. The two problematic transactions---New Order and Payment---rely
on non-monotonic state updates. The former transaction can be modified
to ensure ID uniqueness at the expense of dense ID ordering, while the
latter is an inherently non-monotonic action requiring external
compensation or stronger consistency protocols. We expect that,
especially for read-dominated workloads found in many online services,
HAT guarantees may suffice for many applications. Finally, while TPC-C
is not subject to write skew anomalies, we note that many TPC-E
isolation tests, such as one that requires simultaneously modifying a
product description and its image, are satisfied by HATs.
